\documentclass[11pt,a4paper]{article}

%-------------------------------------------------------
% Encoding and fonts
%-------------------------------------------------------
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\renewcommand{\familydefault}{\sfdefault} % Latin Modern Sans Serif as default
\usepackage{microtype}

%-------------------------------------------------------
% Layout
%-------------------------------------------------------
\usepackage[a4paper,margin=2.4cm]{geometry}
\usepackage{parskip}
\usepackage{enumitem}
\setlist{itemsep=2pt,topsep=3pt}
\usepackage{titlesec}
\titleformat{\section}{\Large\bfseries\sffamily}{\thesection}{0.8em}{}
\titleformat{\subsection}{\large\bfseries\sffamily}{\thesubsection}{0.7em}{}

%-------------------------------------------------------
% Tables, colors, code
%-------------------------------------------------------
\usepackage{array}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{xcolor}
\definecolor{codebg}{HTML}{F4F4F4}
\definecolor{codefr}{HTML}{DADADA}
\definecolor{codekw}{HTML}{1F6FEB}
\definecolor{codest}{HTML}{22863A}
\definecolor{codecm}{HTML}{6A737D}
\definecolor{linkcol}{HTML}{0B5FFF}

\usepackage{listings}
\lstset{
  basicstyle=\ttfamily\small,
  backgroundcolor=\color{codebg},
  frame=single,
  rulecolor=\color{codefr},
  framesep=4pt,
  xleftmargin=6pt,
  xrightmargin=6pt,
  breaklines=true,
  showstringspaces=false,
  keywordstyle=\color{codekw}\bfseries,
  stringstyle=\color{codest},
  commentstyle=\color{codecm}\itshape,
  columns=fullflexible,
  keepspaces=true
}
\lstdefinelanguage{bashplus}{
  morekeywords={up,sudo,git,bash,alias,source,apt,dnf,pacman,zypper,brew,flatpak,snap,systemctl,cron,export},
  morecomment=[l]{\#},
  morestring=[b]",
  morestring=[b]'
}

%-------------------------------------------------------
% Links
%-------------------------------------------------------
\usepackage[colorlinks=true,linkcolor=linkcol,urlcolor=linkcol,citecolor=linkcol]{hyperref}

%-------------------------------------------------------
% Metadata
%-------------------------------------------------------
\title{\textbf{eMerger}\\[4pt]\large One command to update your whole system}
\author{Roberto Antoniello \quad\textbf{\&}\quad Edoardo Ferrari}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
\noindent
eMerger is a cross-platform command line tool that refreshes, upgrades and
cleans a computer with a single command: \texttt{up}. It runs on Linux,
macOS and Windows, auto-detects every package manager present on the host
and invokes them in the right order, returning one unified summary at the
end. This document describes the architecture, features, configuration and
integration points of the current release.
\end{abstract}

\tableofcontents
\vspace{1em}
\hrule
\vspace{1em}

%-------------------------------------------------------
\section{Overview}
%-------------------------------------------------------
eMerger began as a small Bash experiment and has grown into a portable
updater with a shared feature set across the three major desktop and server
platforms. The design follows three principles:

\begin{itemize}
  \item \textbf{One command.} The user types \texttt{up}; everything else
        is automatic detection, ordering and reporting.
  \item \textbf{No runtime bloat.} No daemon, no YAML, no embedded DSL.
        Configuration is a plain shell or PowerShell file that is sourced
        at startup.
  \item \textbf{Use what is there.} eMerger never installs a package
        manager on your behalf; it detects the ones already installed and
        delegates work to them.
\end{itemize}

The project name is a wink at Gentoo's \texttt{emerge}: the tool started
under the name \textit{Updater} and was later renamed to \textit{eMerger}
to reflect its goal of merging the quirks of every supported package
manager behind a uniform interface.

%-------------------------------------------------------
\section{Platforms}
%-------------------------------------------------------

\begin{center}
\begin{tabular}{@{}llll@{}}
\toprule
                  & \textbf{Linux}            & \textbf{macOS}               & \textbf{Windows}        \\
\midrule
Entry point       & \texttt{src/emerger.sh}   & \texttt{src/emerger.sh}      & \texttt{src/emerger.ps1}\\
Shell             & bash 3.2+                 & bash 3.2+ (system)           & PowerShell 5.1+         \\
Setup             & \texttt{./setup.sh}       & \texttt{./setup.sh}          & \texttt{.\textbackslash setup.ps1}   \\
Uninstall         & \texttt{./uninstall.sh}   & \texttt{./uninstall.sh}      & \texttt{.\textbackslash uninstall.ps1}\\
Auto-update       & systemd timer / cron      & cron                         & Task Scheduler          \\
Elevation         & \texttt{sudo}             & \texttt{sudo}                & UAC relaunch            \\
Config dir        & \verb|~/.config/emerger/| & \verb|~/.config/emerger/|    & \verb|%APPDATA%\emerger\|\\
State dir         & \verb|~/.local/state/emerger/| & \verb|~/.local/state/emerger/| & \verb|%LOCALAPPDATA%\emerger\state\|\\
\bottomrule
\end{tabular}
\end{center}

The core flow (detect, upgrade, clean, summarise) is kept at parity across
platforms. Features that depend on operating system primitives, such as
firmware updates or filesystem snapshots, are gated to the platforms that
support them and are marked as such throughout this document.

%-------------------------------------------------------
\section{Installation}
%-------------------------------------------------------

\subsection{Linux and macOS}

\begin{lstlisting}[language=bashplus]
git clone https://github.com/MasterCruelty/eMerger
cd eMerger
./setup.sh
# open a new shell (or: source ~/.bashrc)
up --help
\end{lstlisting}

\texttt{setup.sh} performs the following actions:

\begin{enumerate}
  \item Adds \texttt{alias up='bash /path/to/eMerger/src/emerger.sh'}
        to the appropriate shell rc file (bash, zsh or fish).
  \item Marks \texttt{src/emerger.sh} as executable.
  \item Installs shell completions for bash, zsh and fish.
  \item Scaffolds \verb|~/.config/emerger/| with the default
        \texttt{config.sh}, \texttt{ignore.list}, \texttt{hooks/} and
        \texttt{profiles.d/} layout.
\end{enumerate}

On macOS the alias is written to \texttt{\string~/.zshrc} first, since zsh
is the default interactive shell from Catalina onwards. When Homebrew is
present, completions are installed into the Homebrew prefix so they are
loaded by the brew-managed shell environment.

\subsection{Windows}

\begin{lstlisting}[language=bashplus]
git clone https://github.com/MasterCruelty/eMerger
cd eMerger
.\setup.ps1
# open a new PowerShell window (or: . $PROFILE)
up --help
\end{lstlisting}

\texttt{setup.ps1} does not require administrator privileges. It lifts
the execution policy for the current user to \texttt{RemoteSigned} when
needed, defines a \texttt{function up} inside
\texttt{\$PROFILE.CurrentUserAllHosts} so that the command is available
from every PowerShell host, and scaffolds \verb|%APPDATA%\emerger\|. When a
package manager invocation later needs elevation, eMerger relaunches
itself under UAC.

\subsection{Manual install}

When a packaged setup is undesirable, pointing an alias at the entry point
is enough:

\begin{lstlisting}[language=bashplus]
alias up='bash /absolute/path/to/eMerger/src/emerger.sh'
\end{lstlisting}

\begin{lstlisting}[language=bashplus]
function up { & "C:\path\to\eMerger\src\emerger.ps1" @args }
\end{lstlisting}

\subsection{Uninstall}

Running \texttt{./uninstall.sh} (Unix) or \texttt{.\textbackslash uninstall.ps1}
(Windows) removes the shell alias, the cron job or systemd timer, and the
scheduled task. Configuration and state directories are preserved so that
logs, profiles and hooks survive a reinstall; remove them manually to
start from a clean slate.

%-------------------------------------------------------
\section{Self-update}
%-------------------------------------------------------
Three equivalent forms are accepted:

\begin{lstlisting}[language=bashplus]
up -up
up --self-update
./update.sh        # .\update.ps1 on Windows
\end{lstlisting}

Internally a \texttt{git pull --ff-only} is performed inside the clone and
the new commit range is printed. Non fast-forward updates are refused to
guarantee that local changes cannot silently disappear.

%-------------------------------------------------------
\section{Execution flow}
%-------------------------------------------------------
A full run of \texttt{up} performs, in order:

\begin{enumerate}
  \item Load \texttt{config.sh} (or \texttt{config.ps1}) and any
        \texttt{-\/-profile}.
  \item Parse CLI flags; CLI always wins over config and profile defaults.
  \item Acquire an exclusive lock (\verb|/tmp/emerger.lock| on Unix).
  \item Print the logo, OS info line and timestamp.
  \item Warn on low battery and on low free disk space.
  \item Cache \texttt{sudo} credentials on Unix, or relaunch elevated on
        Windows, only when a detected manager actually needs it.
  \item Snapshot the list of installed packages for the post-run diff.
  \item Execute \texttt{pre.d} hooks.
  \item For each detected manager, run \textit{refresh}, \textit{upgrade},
        \textit{clean}.
  \item Optionally clear user cache and trash.
  \item Execute \texttt{post.d} hooks.
  \item Compute the installed-packages diff.
  \item Render the summary, the reboot advisory and an optional desktop
        notification.
\end{enumerate}

%-------------------------------------------------------
\section{Command line reference}
%-------------------------------------------------------

The authoritative reference is \texttt{up -\/-help}. The most frequently
used flags are summarised below.

\begin{center}
\begin{longtable}{@{}p{0.28\linewidth} p{0.52\linewidth} p{0.14\linewidth}@{}}
\toprule
\textbf{Flag} & \textbf{Meaning} & \textbf{Platforms} \\
\midrule \endhead
\texttt{-n}, \texttt{-\/-dry-run}        & Preview without running       & all \\
\texttt{-v}, \texttt{-\/-verbose}        & Stream manager output live    & all \\
\texttt{-q} / \texttt{-qq} / \texttt{-qqq} & Progressive quiet levels   & all \\
\texttt{-y}, \texttt{-\/-yes}            & Assume yes on prompts         & all \\
\texttt{-i}, \texttt{-\/-interactive}    & TUI menu (gum / whiptail)     & Unix \\
\texttt{-\/-security}                    & Security-only updates         & supported managers \\
\texttt{-\/-dev}                         & Include dev toolchains        & all \\
\texttt{-\/-firmware}                    & \texttt{fwupdmgr} refresh+upd & Linux \\
\texttt{-\/-parallel}                    & Run user-space in parallel    & Unix \\
\texttt{-\/-profile NAME}                & Load profile                  & all \\
\texttt{-\/-snapshot}                    & snapper / timeshift / btrfs   & Linux \\
\texttt{-\/-refresh-mirrors}             & Re-rank distribution mirrors  & Linux \\
\texttt{-\/-resume}                      & Skip completed managers       & Unix \\
\texttt{-\/-reboot}                      & Reboot if required            & all \\
\texttt{-\/-changed}                     & Show package diff             & Unix \\
\texttt{-\/-changelog PKG}               & Upstream changelog for PKG    & Unix \\
\texttt{-\/-report FILE}                 & Export Markdown report        & Unix \\
\texttt{-\/-doctor}                      & Health check                  & all \\
\texttt{-\/-history}                     & Recent runs                   & all \\
\texttt{-\/-errors}, \texttt{-err}       & Tail of logged errors         & all \\
\texttt{-\/-json}                        & Machine-readable summary      & all \\
\texttt{-\/-reboot-exit}                 & Exit 4 if reboot is required  & all \\
\texttt{-\/-rollback}                    & Revert last snapshot          & Linux \\
\texttt{-\/-download-only}, \texttt{-\/-offline} & Prefetch, don't install & Unix \\
\texttt{-\/-only LIST}                   & Only these managers           & all \\
\texttt{-\/-except LIST}                 & Skip these managers           & all \\
\texttt{-\/-metrics FILE}                & Prometheus textfile export    & all \\
\texttt{-up}                             & Self-update eMerger           & all \\
\texttt{-au}                             & Install weekly auto-update    & all \\
\bottomrule
\end{longtable}
\end{center}

\subsection{Combining flags}
Flags are independent tokens and can be combined freely:

\begin{lstlisting}[language=bashplus]
up -n -v                    # preview a full run, live stream
up -y -q --security         # unattended security-only, minimal output
up --dev --parallel -v      # include dev toolchains, parallel user-space
up --snapshot --reboot -y   # snapshot first, reboot at the end if needed
up --profile server --resume
\end{lstlisting}

Single-letter short flags in the set \texttt{\{h V n v q y i w\}} also
bundle: \texttt{up -nv} equals \texttt{up -n -v}. Compound short flags
(\texttt{-nl}, \texttt{-ni}, \texttt{-qq}, \texttt{-up}, \texttt{-err},
\dots) and long flags are never bundled.

%-------------------------------------------------------
\section{Features in depth}
%-------------------------------------------------------

\subsection{Dry-run and verbose}
\texttt{-n} prints every command that would be executed without actually
running it or asking for elevation, which makes it safe to run in a
terminal that does not own sudo credentials. \texttt{-v} streams the
output of each package manager live instead of buffering it into the
summary. The two combine naturally.

\subsection{Quiet levels}
\texttt{-q} hides informational output, \texttt{-qq} keeps only the step
titles and the final one-line summary, \texttt{-qqq} produces no output
at all and communicates only through the exit code.

\subsection{Security-only updates}
When \texttt{-\/-security} is passed, each supporting manager is asked to
limit the upgrade set to security advisories: \texttt{apt} via
\texttt{unattended-upgrade}, \texttt{dnf} with \texttt{-\/-security},
\texttt{zypper patch -\/-category security}, macOS via
\texttt{softwareupdate -\/-recommended} and Windows via
\texttt{PSWindowsUpdate} severity filtering. Managers that do not
distinguish between security and non-security updates ignore the flag.

\subsection{Dev toolchains}
\texttt{-\/-dev} enables a separate track that runs \texttt{rustup}
updates, \texttt{cargo install-update -a}, \texttt{npm update -g},
\texttt{pnpm -g update}, \texttt{pip} user site upgrades and
\texttt{gem update}. These run under the invoking user, never under
sudo.

\subsection{Firmware (Linux)}
\texttt{-\/-firmware} runs
\texttt{fwupdmgr refresh \&\& fwupdmgr update -y -\/-no-reboot-check}.
On macOS firmware is handled by \texttt{softwareupdate}; on Windows it
is delegated to vendor tooling (Dell Command Update, Lenovo Vantage,
etc.) and is out of scope.

\subsection{Parallel mode}
\texttt{-\/-parallel} runs user-space managers that do not touch the root
filesystem concurrently: \texttt{flatpak}, \texttt{snap}, \texttt{brew},
\texttt{mas} and dev toolchains. System managers stay serial to avoid
lock contention. The Windows side is currently always serial.

\subsection{Snapshots and rollback (Linux)}
\texttt{-\/-snapshot} tries, in order, \texttt{snapper},
\texttt{timeshift} and raw \texttt{btrfs subvolume snapshot}.
\texttt{-\/-rollback} reverts to the most recent eMerger-created snapshot;
\texttt{timeshift} is always invoked interactively because a silent
rollback is too destructive to automate.

\subsection{Mirrors refresh (Linux)}
\texttt{-\/-refresh-mirrors} re-ranks the repository mirrors before the
upgrade: \texttt{reflector -\/-latest 20 -\/-sort rate} on Arch,
\texttt{netselect-apt} on Debian and Ubuntu, and a no-op on Fedora
where \texttt{fastestmirror} already handles ranking at the DNF level.

\subsection{Reboot handling}
On Unix, eMerger inspects \texttt{/var/run/reboot-required} and
\texttt{needs-restarting -r}. On Windows it reads the
\texttt{CBS RebootPending}, \texttt{WindowsUpdate\textbackslash RebootRequired}
and \texttt{PendingFileRenameOperations} registry entries. The summary
prints a reboot advisory; \texttt{-\/-reboot} actually reboots and
\texttt{-\/-reboot-exit} exits with code 4 so that an orchestrator can
decide.

\subsection{Resume}
When a run is interrupted, a cursor is written to
\verb|~/.local/state/emerger/resume|. A subsequent \texttt{up -\/-resume}
skips every manager that completed successfully, so repeated failures do
not force a full retry. Unix only.

\subsection{Package diff and changelog}
Each run records the installed package set before and after. \texttt{up
-\/-changed} prints the diff with \texttt{+}, \texttt{-} and \texttt{\~}
markers for added, removed and upgraded packages. \texttt{up -\/-changelog
PKG} dispatches to \texttt{apt changelog}, \texttt{dnf changelog},
\texttt{pacman -Qi} or \texttt{brew log} depending on the host.

\subsection{Reports, history and errors}
\texttt{-\/-report FILE} exports a Markdown report of the last run.
\texttt{-\/-history} prints a table of the most recent runs.
\texttt{-\/-errors} tails \texttt{ERROR} lines from the log.

\subsection{Doctor}
\texttt{-\/-doctor} audits the environment: shell or PowerShell version,
sudo or admin state, disk space, network reachability, state directory
writability, per-manager native health checks
(\texttt{dpkg -\/-audit}, \texttt{pacman -Dk}, \texttt{brew doctor}),
pending reboot flags and, on Windows, the current execution policy. A
non-zero exit code is returned when any issue is detected.

%-------------------------------------------------------
\section{Configuration}
%-------------------------------------------------------

\subsection{Config file}

\textbf{Unix} (\verb|~/.config/emerger/config.sh|, sourced before argument
parsing):

\begin{lstlisting}[language=bashplus]
ARG_DEV=1                  # always include dev toolchains
ARG_WEATHER=1              # show weather widget
DISK_MIN_FREE_MB=2048      # require at least 2 GB free on /
QUIET_HOURS="23:00-07:00"  # skip scheduled runs in this window
RETRY_MAX=3                # transient-error retry count
\end{lstlisting}

\textbf{Windows} (\verb|%APPDATA%\emerger\config.ps1|, dot-sourced):

\begin{lstlisting}[language=bashplus]
$script:ArgsGlobal.Dev      = $true
$script:ArgsGlobal.Security = $true
$script:ArgsGlobal.NoTrash  = $true
\end{lstlisting}

\subsection{Profiles}
Profiles are named configuration snippets loaded through
\texttt{-\/-profile NAME}. The shipped set lives under
\texttt{share/profiles/} and includes \texttt{work}, \texttt{home},
\texttt{server} and \texttt{safe}. User overrides live in
\verb|~/.config/emerger/profiles.d/| on Unix and
\verb|%APPDATA%\emerger\profiles.d\| on Windows, and always shadow the
shipped defaults.

\subsection{Hooks}
Executable scripts in \texttt{hooks/pre.d/} and \texttt{hooks/post.d/}
run before and after the update flow, in alphabetical order. A failing
hook emits a warning but does not abort the run, so hooks are safe by
default. Unix hooks are \texttt{*.sh} under bash; Windows hooks are
\texttt{*.ps1} dot-sourced under PowerShell.

\begin{lstlisting}[language=bashplus]
# ~/.config/emerger/hooks/pre.d/10-backup-dotfiles.sh
#!/usr/bin/env bash
rsync -a ~/.config/ ~/backups/dotfiles/
\end{lstlisting}

\subsection{Ignore list (Linux)}
\verb|~/.config/emerger/ignore.list| accepts one package per line, with
\texttt{\#} comments. \texttt{pacman} honours it natively via
\texttt{-\/-ignore=}. For \texttt{apt}, \texttt{dnf} and \texttt{zypper}
the file is advisory: the user must still hold the packages through the
manager's native command (\texttt{apt-mark hold},
\texttt{dnf versionlock add}, \texttt{zypper al}).

\subsection{Quiet hours}
\texttt{QUIET\_HOURS="HH:MM-HH:MM"} in \texttt{config.sh} makes scheduled
runs that start inside the window exit immediately when \texttt{-y} is
set, i.e. the ones coming from the timer. Interactive invocations are
never blocked.

\subsection{Manager plugins}
Dropping a bash script in \verb|~/.config/emerger/managers.d/<name>.sh|
registers a custom manager without touching the repository. A minimal
plugin declares \texttt{pm\_<slug>\_detect} and \texttt{pm\_<slug>\_run};
optional callbacks expose whether the manager needs sudo, can run in
parallel, is gated by \texttt{-\/-dev} and which icon to print. Plugins
honour \texttt{-\/-only}, \texttt{-\/-except}, \texttt{-\/-parallel},
\texttt{-\/-dev}, the detection cache, hooks and the summary. A
copy-pasteable template is provided in
\texttt{share/plugins/example.sh}. Linux and macOS only.

%-------------------------------------------------------
\section{Integration}
%-------------------------------------------------------

\subsection{JSON output}
\texttt{up -\/-json} emits a single-line JSON object on stdout, with the
logo, info line and summary box suppressed. The same schema is what
\texttt{history.jsonl} stores one line at a time:

\begin{lstlisting}[language=bashplus]
{"ts":"2026-04-14T07:24:31Z","duration":42,"freed_kb":15360,
 "errors":0,"reboot":0,
 "managers":[{"name":"apt","result":"ok"},{"name":"flatpak","result":"ok"}]}
\end{lstlisting}

\subsection{Prometheus metrics}
\texttt{up -\/-metrics FILE} reads the most recent entry of
\texttt{history.jsonl} and writes a Prometheus textfile-collector
snapshot. Exposed gauges include
\texttt{emerger\_last\_run\_timestamp\_seconds},
\texttt{emerger\_last\_run\_duration\_seconds},
\texttt{emerger\_last\_run\_freed\_bytes},
\texttt{emerger\_last\_run\_errors},
\texttt{emerger\_reboot\_required} and one
\texttt{emerger\_manager\_ok\{manager="..."\}} per manager of the last
run. The flag does not trigger an update run; invoke it from a
\texttt{post.d} hook or from the scheduler after \texttt{up} completes.

\subsection{Download-only and offline}
\texttt{-\/-download-only} (alias \texttt{-\/-offline}) refreshes the
package indexes and downloads the pending upgrade set but does not
install it. Supported natively on \texttt{pacman} (\texttt{-Syuw}),
\texttt{apt} (\texttt{-\/-download-only}), \texttt{dnf}
(\texttt{-\/-downloadonly}) and \texttt{zypper}
(\texttt{update -\/-download-only}). This is the safe prefetch option
for metered connections and for pre-staging maintenance windows.

\subsection{Manager filtering}
\texttt{-\/-only LIST} restricts the run to the listed comma-separated
managers; \texttt{-\/-except LIST} excludes them. Filtering is applied
after detection and after \texttt{-\/-dev} / \texttt{-\/-firmware}
gating, and \texttt{-\/-except} always wins over \texttt{-\/-only} when
both mention the same name.

%-------------------------------------------------------
\section{Auto-update (unattended)}
%-------------------------------------------------------
\texttt{up -au} installs a weekly unattended run. On Linux the preferred
backend is a systemd user timer
(\verb|~/.config/systemd/user/emerger.{service,timer}|); cron is used as
fallback. On macOS cron is the default and \texttt{launchd} can be used
manually. On Windows a \texttt{Register-ScheduledTask -TaskName eMerger}
is created. The trigger fires weekly on Sunday at 10:00 with a
$\pm$ one-hour randomised delay, and always uses
\texttt{-y -q -nl -ni}. Pair with \texttt{QUIET\_HOURS} to keep the
machine quiet at night.

%-------------------------------------------------------
\section{Files and paths}
%-------------------------------------------------------

\paragraph{Linux and macOS}
\begin{center}
\begin{tabular}{@{}p{0.45\linewidth} p{0.45\linewidth}@{}}
\toprule
\textbf{Path} & \textbf{Purpose} \\
\midrule
\verb|~/.config/emerger/config.sh|        & User defaults \\
\verb|~/.config/emerger/profiles.d/|      & User profiles \\
\verb|~/.config/emerger/hooks/{pre,post}.d/| & Hook scripts \\
\verb|~/.config/emerger/ignore.list|      & Ignore list (pacman native) \\
\verb|~/.config/emerger/managers.d/*.sh|  & User-defined manager plugins \\
\verb|~/.cache/emerger/detected|          & Detection cache \\
\verb|~/.local/state/emerger/emerger.log| & Log (rotated at 2000 lines) \\
\verb|~/.local/state/emerger/history.jsonl| & One JSON record per run \\
\verb|~/.local/state/emerger/resume|      & Resume cursor \\
\verb|/tmp/emerger.lock|                  & Global lock (\texttt{flock}) \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Windows}
\begin{center}
\begin{tabular}{@{}p{0.45\linewidth} p{0.45\linewidth}@{}}
\toprule
\textbf{Path} & \textbf{Purpose} \\
\midrule
\verb|%APPDATA%\emerger\config.ps1|       & User defaults \\
\verb|%APPDATA%\emerger\profiles.d\|      & User profiles \\
\verb|%APPDATA%\emerger\hooks\{pre,post}.d\| & Hook scripts \\
\verb|%LOCALAPPDATA%\emerger\cache\|      & Detection cache \\
\verb|%LOCALAPPDATA%\emerger\state\emerger.log|   & Log \\
\verb|%LOCALAPPDATA%\emerger\state\history.jsonl| & Run history \\
\bottomrule
\end{tabular}
\end{center}

%-------------------------------------------------------
\section{Source tree}
%-------------------------------------------------------
The Bash and PowerShell implementations share the same layering:
argument parsing, UI, logging, locking and per-manager dispatch are each
isolated in their own file.

\paragraph{Bash (\texttt{src/lib/})}
\texttt{args.sh}, \texttt{changelog.sh}, \texttt{clean.sh},
\texttt{diff.sh}, \texttt{disk.sh}, \texttt{doctor.sh},
\texttt{estimate.sh}, \texttt{hooks.sh}, \texttt{ignore.sh},
\texttt{lock.sh}, \texttt{log.sh}, \texttt{metrics.sh},
\texttt{mirrors.sh}, \texttt{notify.sh}, \texttt{packages.sh},
\texttt{plugins.sh}, \texttt{profiles.sh}, \texttt{progress.sh},
\texttt{reboot.sh}, \texttt{report.sh}, \texttt{resume.sh},
\texttt{retry.sh}, \texttt{run.sh}, \texttt{snapshot.sh},
\texttt{summary.sh}, \texttt{sys.sh}, \texttt{tui.sh},
\texttt{ui.sh}, \texttt{update.sh}, \texttt{wizard.sh},
\texttt{help.txt}.

\paragraph{PowerShell (\texttt{src/pslib/})}
\texttt{Args.ps1}, \texttt{Clean.ps1}, \texttt{Doctor.ps1},
\texttt{Hooks.ps1}, \texttt{Log.ps1}, \texttt{Notify.ps1},
\texttt{Packages.ps1}, \texttt{Profiles.ps1}, \texttt{Summary.ps1},
\texttt{Sys.ps1}, \texttt{UI.ps1}, \texttt{Update.ps1},
\texttt{Help.txt}.

Adding a new package manager means extending
\texttt{src/lib/packages.sh} on the Unix side or
\texttt{src/pslib/Packages.ps1} on the Windows side; both are
table-driven dispatchers.

%-------------------------------------------------------
\section{Supported package managers}
%-------------------------------------------------------
\begin{itemize}
  \item \textbf{Linux, system (sudo):} \texttt{pacman}, \texttt{apt},
        \texttt{apt-get}, \texttt{dnf}, \texttt{yum}, \texttt{zypper},
        \texttt{xbps}, \texttt{apk}, \texttt{eopkg}, \texttt{emerge},
        \texttt{nixos-rebuild}, \texttt{fwupdmgr}, \texttt{snap}.
  \item \textbf{Linux, AUR (no sudo):} \texttt{yay}, \texttt{paru}.
  \item \textbf{Linux, user-space:} \texttt{flatpak}, \texttt{nix-env}.
  \item \textbf{macOS:} \texttt{softwareupdate}, \texttt{brew},
        \texttt{brew --cask}, \texttt{mas}.
  \item \textbf{Windows:} \texttt{winget}, \texttt{scoop},
        \texttt{choco}, \texttt{PSWindowsUpdate}, \texttt{wsl --update}.
  \item \textbf{Dev toolchains (all platforms, opt-in):} \texttt{rustup},
        \texttt{cargo install-update}, \texttt{npm}, \texttt{pnpm},
        \texttt{pip}, \texttt{gem}.
\end{itemize}

%-------------------------------------------------------
\section{Exit codes}
%-------------------------------------------------------
\begin{center}
\begin{tabular}{@{}cl@{}}
\toprule
\textbf{Code} & \textbf{Meaning} \\
\midrule
0 & Success \\
1 & Runtime failure (sudo, lock, disk, interrupted) \\
2 & Argument parsing error \\
3 & One or more package managers returned non-zero \\
4 & Reboot required (only with \texttt{-\/-reboot-exit}) \\
\bottomrule
\end{tabular}
\end{center}

%-------------------------------------------------------
\section{Troubleshooting}
%-------------------------------------------------------
\textbf{Another eMerger run is in progress.} A stale \texttt{flock} on
\verb|/tmp/emerger.lock|. Inspect \texttt{ps} for stragglers, then
remove the lock file.

\textbf{A manager fails with exit code 3.} Run \texttt{up -\/-errors}
to tail the log, then \texttt{up -\/-doctor} to inspect the environment.
Most failures come from a manager's own health state (for example
\texttt{dpkg} being interrupted); run the native repair command before
retrying.

\textbf{Nothing happens on scheduled runs.} Check
\texttt{systemctl -\/-user status emerger.timer} or the Windows Task
Scheduler entry. On Unix the scheduled run always uses
\texttt{-y -q -nl -ni}; enable \texttt{-v} in the unit for one cycle to
see the output.

%-------------------------------------------------------
\section{Contributing}
%-------------------------------------------------------
Pull requests and issues are welcome on
\href{https://github.com/MasterCruelty/eMerger}{GitHub}. Before opening
a PR, run the test suite in \texttt{tests/} on the platform you are
touching, and make sure the feature is covered on both the Bash and the
PowerShell sides when it makes sense for both platforms.

\vspace{1em}
\hrule
\vspace{0.5em}
\noindent\textit{If your question is not answered here, open an issue on
GitHub: we would rather improve the documentation than answer the same
question twice.}

\end{document}
